---
title: "ManyFaces pilot ratings"
subtitle: "Rater demographics and interrater agreement"
format: 
  html:
    code-fold: true
    df-print: paged
    embed-resources: true
    fig-height: 5
    fig-width: 8
    toc: true
execute: 
  error: true
---

```{r}
#| message: false
library(psych) # ICC, alpha, omega
library(furrr) # run code parallel
library(tidyverse)

# Load custom functions
source("data/custom_functions.R")
```

## Load data
```{r}
# Rating data
data_exp <- read_csv("data/manyfaces-pilot-exp_cleaned.csv",
                     show_col_types = FALSE)

# Questionnaire data
data_qu <- read_csv("data/manyfaces-pilot-quest_cleaned.csv",
                    show_col_types = FALSE)
```

```{r}
data_trait <- data_exp |>
  filter(exp == "attractive" | exp == "dominant"  | exp == "trustworthy"  | exp == "gender-typical"  | exp == "memorable") |>
  mutate(exp = factor(exp, levels = c("attractive", "dominant", "trustworthy", "gender-typical", "memorable"))) |>
  mutate(dv = as.numeric(dv))

data_trait_unstd <- data_exp |>
  filter(grepl("unstd", exp)) |>
  mutate(dv = as.numeric(dv))

data_emo_int <- data_exp |>
  filter(exp == "anger" | exp == "happiness"  | exp == "disgust"  | exp == "surprise"  | exp == "sadness" | exp == "fear") |>
  mutate(dv = as.numeric(dv))

data_emo_cat <- data_exp |> 
  filter(grepl("^em", exp))
```

## Rater demographics

Demographic data is available for 1685 of the total of 1909 raters.

### Age and gender

```{r}
data_qu |> 
  ggplot(aes(x = age, fill = gender)) +
  geom_histogram(binwidth = 1) +
  scale_fill_manual(values = c("hotpink", "lightblue", "orchid", "gray"))
```

### Residence

```{r}
count(data_qu, residence, sort = TRUE)
```

### Ethnicity

```{r}
data_qu |>
  count(ethnicity_rec, sort = TRUE)
```

### Devices

```{r}
count(data_qu, device, sort = TRUE)
```

## Interrater agreement

### Standardised trait ratings

#### ICC

**Observed across experiments**
```{r}
data_trait |> 
  group_by(exp) |> 
  group_map(calc_icc) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), ~ round(.x, 2)))
```

**"Corridors of stability"**
At what sample size (if at all) do ratings reach acceptable reliability of .75 to .90?
```{r}
#| warning: false
#| message: false

# Set seed for reproducibility
set.seed(123)

# Set a stable parallel plan
future::plan(multisession, workers = 3)

# Get unique experiments
experiments <- unique(data_trait$exp)

# Run the function in parallel across experiments
corridor_icc_list <- furrr::future_map(
  experiments,
  function(exp_name) {
    data_subset <- data_trait |> filter(exp == exp_name)
    run_or_load_icc(
      exp_name = exp_name,
      data = data_subset,
      n_raters_seq = seq(10, 100, 10),
      n_iter = 50
    )
  },
  .options = furrr::furrr_options(seed = TRUE)
)

# Combine results
icc_corridor <- list_rbind(corridor_icc_list)

# Reset plan to sequential
future::plan(sequential)
```

```{r}
icc2k_corridor_summary <- icc_corridor |> 
  group_by(experiment, n_raters_sampled) |> 
  summarise(
    ICC2_k_median = median(`ICC(2,k)`, na.rm = TRUE),
    ICC2_k_low = quantile(`ICC(2,k)`, 0.025, na.rm = TRUE),
    ICC2_k_high = quantile(`ICC(2,k)`, 0.975, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
icc2k_corridor_summary |> 
  ggplot(aes(x = n_raters_sampled)) +
  geom_ribbon(aes(ymin = ICC2_k_low, ymax = ICC2_k_high), 
              alpha = 0.15, fill = "grey60") +
  # Median ICC(2,k)
  geom_line(aes(y = ICC2_k_median), linewidth = 0.8, colour = "black") +
  # Thresholds: 0.75 grey, 0.9 black
  geom_hline(yintercept = 0.75, linetype = "dashed", colour = "grey40") +
  geom_hline(yintercept = 0.9, linetype = "dashed", colour = "black") +
  facet_wrap(~ experiment) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Number of raters sampled", 
    y = "ICC(2,k)") +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 12),
    plot.subtitle = element_text(size = 11)
  )
```

#### Alpha & omega

**Observed across experiments**
Note on warnings:
 * "Some items were negatively correlated...": potentially reflects true inter-rater disagreement; reversing raters not appropriate in this context
 * "Omega_h... not meaningful": omega_h is undefined when computing a one-factor solution; omega_t remains valid and is the relevant statistic reported.

```{r}
#| message: false
data_trait |> 
  group_by(exp) |> 
  group_map(purrr::partial(calc_alpha_omega)) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

### Unstandardised trait ratings

#### ICC
```{r}
data_trait_unstd |> 
  group_by(exp) |> 
  group_map(calc_icc, check_dropped_raters = FALSE) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

#### Alpha & omega
```{r}
#| message: false
data_trait_unstd |> 
  group_by(exp) |> 
  group_map(calc_alpha_omega) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

### Emotion intensity ratings

#### ICC
```{r}
data_emo_int |> 
  group_by(exp) |> 
  group_map(calc_icc) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```

#### Alpha & omega
```{r}
#| warning: false
#| message: false
data_emo_int |> 
  group_by(exp) |> 
  group_map(calc_alpha_omega) |> 
  bind_rows() |> 
  mutate(across(where(is.numeric), \(x) round(x, 2)))
```


## Point of stability (Hehman et al., 2018)

### Standardised trait ratings

```{r}
set.seed(123)

system.time({
  stability_stats_traits <- calc_stability_stats(data = data_trait,
                                               N = 100,
                                               iterations = 300,
                                               ci_interval = 0.95, ci_method = "percentile",
                                               cos_threshold = 0.5,
                                               save_means = FALSE, # only save if actually needed (huge)
                                               col_map = list(trait = "exp",
                                                              stim_id = "trial_name",
                                                              rating = "dv")) 
})

```

```{r}
summary_ci_traits <- stability_stats_traits$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```


Plot
```{r}
stability_stats_traits$pos$label <- paste0("Point of stability\n(N = ", stability_stats_traits$pos$pos, ")")

summary_ci_traits |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_traits$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_traits$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```

Sanity check with Hehman et al. (20180) code

```{r}
source("data/resampling_hehmanetal.R")
```

```{r}
#| warning: false
#| message: false
# Trustworthy
trustworthy <- data_trait |> 
  filter(exp == "trustworthy")

trustworthy_rs <-RSSample(Rating = trustworthy$dv,
                          Stimulus = factor(trustworthy$trial_name))

# Our code N=37; Hehman et al. (2018) N=37
RSGetPointOfStability(trustworthy_rs$ci, 0.5, inarow = 1)

RSFigure(
  trustworthy_rs$data,
  trustworthy_rs$ci,
  threshold = .5,
  title = "",
  figPNGname = "",
  inarow = 1
)
```

```{r}
#| warning: false
#| message: false
# Memorable
memorable <- data_trait |> 
  filter(exp == "memorable")

memorable_rs <-RSSample(Rating = memorable$dv,
                        Stimulus = factor(memorable$trial_name))

# Our code N=45; Hehman et al. (2018) N=45
RSGetPointOfStability(memorable_rs$ci, 0.5, inarow = 1)

RSFigure(
  memorable_rs$data,
  memorable_rs$ci,
  threshold = .5,
  title = "",
  figPNGname = "",
  inarow = 1
)
```

### Unstandardised trait ratings

```{r}
set.seed(123)

stability_stats_traits_unstd <- calc_stability_stats(data = data_trait_unstd,
                                                     N = 100,
                                                     iterations = 300,
                                                     ci_interval = 0.95, ci_method = "percentile",
                                                     cos_threshold = 0.5,
                                                     save_means = FALSE,
                                                     col_map = list(trait = "exp",
                                                                    stim_id = "trial_name",
                                                                    rating = "dv")) 

summary_ci_traits_undstd <- stability_stats_traits_unstd$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```

Plot
```{r}
# POS labels
stability_stats_traits_unstd$pos$label <- paste0("Point of stability\n(N = ", stability_stats_traits_unstd$pos$pos, ")")

summary_ci_traits_undstd |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_traits_unstd$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_traits_unstd$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```

### Emotion intensity ratings
```{r}
set.seed(123)

stability_stats_emo_int <- calc_stability_stats(data = data_emo_int,
                                                N = 100,
                                                iterations = 300,
                                                ci_interval = 0.95, ci_method = "percentile",
                                                cos_threshold = 0.5,
                                                save_means = FALSE,
                                                col_map = list(trait = "exp",
                                                               stim_id = "trial_name",
                                                               rating = "dv")) 

summary_ci_emo_int <- stability_stats_emo_int$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```

Plot
```{r}
# POS labels
stability_stats_emo_int$pos$label <- paste0("Point of stability\n(N = ", stability_stats_emo_int$pos$pos, ")")

summary_ci_emo_int |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_emo_int$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_emo_int$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```