---
title: "ManyFaces Pilot Ratings Data Prep"
format: 
  html:
    code-fold: true
    df-print: paged
    embed-resources: true
    fig-height: 5
    fig-width: 8
    toc: true
execute: 
  error: true
---

```{r}
#| message: false
library(tidyverse)
```


## Project Structure

Read in the full project structure from the project file to map experiment names to numbers. 

```{r}
proj <- jsonlite::read_json("data/project_1136_structure.json")

exp_data <- purrr::map_df(proj, \(comp) {
  if (comp$component_type == "exp") {
     data.frame(
       exp_id = comp$id,
       name = comp$name,
       res_name = comp$res_name,
       instructions = comp$instructions,
       question = comp$question,
       exptype = comp$exptype,
       trial_order = comp$trial_order,
       total_stim = comp$total_stim,
       random_stim = comp$random_stim,
       trials = length(comp$trial),
       stim = length(comp$stim)
     )
  } else {
    NULL
  }
}) |>
  mutate(exp = sub("ManyFaces? Pilot Ratings: ", "", res_name) |> trimws())


trial_data <- purrr::map_df(proj, \(comp) {
  if (comp$component_type == "exp") {
    purrr::map_df(comp$trial, \(trial) {
      data.frame(
        exp_id = comp$id,
        n = trial$trial_n,
        name = trial$name, 
        img_id = trial$center_img,
        img_path = comp$stimuli[[as.character(trial$center_img)]]
      )
    })
  } else {
    NULL
  }
}) |>
  mutate(name = sub("^(manyfaces|attention_checks)/", "", name))
```

```{r}
exp_data |>
  select(exp_id, exp, question, trials) |>
  arrange(exp)
```


## Data Prep

### Raw Data

This workflow requires the data-raw directory, which is not shared on github. 

This is the SQL for downloading the data from Experimentum. We need to download in chunks of 50000 rows to avoid file download limits on the site (not needed if downloading directly from SQL).

``` sql
SELECT 
  session.id as session_id, project_id, exp.res_name as exp_name, exp_id, 
  session.user_id, user.sex as user_sex, user.status as user_status,
  ROUND(DATEDIFF(ed.dt, REPLACE(birthday, "-00","-01"))/365.25, 1) AS user_age,
  trial.name as trial_name,
  trial_n,
  `order`,
  dv,
  rt,
  ed.side,
  ed.dt
FROM session 
  LEFT JOIN user USING (user_id)
  LEFT JOIN exp_data AS ed ON ed.session_id = session.id
  LEFT JOIN exp ON exp.id = ed.exp_id
  LEFT JOIN trial USING (exp_id, trial_n)
WHERE session.project_id = 1136
  AND user.status IN ("guest", "registered")
  AND exp_id IN (1384, 1400, 1399, 1398, 1401, 1402, 1403, 
                 1404, 1405, 1397, 1390, 1389, 1388, 1387, 
                 1386, 1385, 1382, 1381, 1380, 1379, 1377)
LIMIT 50000
OFFSET 0
```

```{r}
#| eval: false

# combine multiple downloads into one file
exp_raw <- list.files("data-raw/exp", full.names = TRUE) |> 
  read_csv(show_col_types = FALSE) |>
  unique() |>
  filter(user_status %in% c("guest", "registered"))

write_csv(exp_raw, paste0("data-raw/ManyFaces-Pilot-Ratings-exps_", Sys.Date(), ".csv"))

# get most recent files
exp_file <- list.files("data-raw", "ManyFaces-Pilot-Ratings-exps",
                       full.names = TRUE) |>
  sort(decreasing = TRUE) |>
  pluck(1)

exp_raw <- read_csv(exp_file, show_col_types = FALSE) |>
  filter(user_status %in% c("guest", "registered")) |>
  unique()

# get most recent files
quest_file <- list.files("data-raw", "ManyFaces-Pilot-Ratings-quests", 
                       full.names = TRUE) |>
  sort(decreasing = TRUE) |>
  pluck(1)

quest_raw <- read_csv(quest_file, show_col_types = FALSE) |>
  filter(user_status %in% c("guest", "registered")) |>
  unique()

# write to the data directory
write_csv(exp_raw, "data/manyfaces-pilot-exp.csv")
write_csv(quest_raw, "data/manyfaces-pilot-quest.csv")
```


### Reshape Data

```{r}
exp_raw <- read_csv("data/manyfaces-pilot-exp.csv", show_col_types = FALSE)
quest_raw <- read_csv("data/manyfaces-pilot-quest.csv", show_col_types = FALSE)
```


```{r}
ed <- exp_data |> select(exp_id, exp)

exp_long <- exp_raw |>
  select(session_id, exp_id, trial_name, dv, rt, dt) |>
  unique() |>
  mutate(trial_name = sub("^(manyfaces|attention_checks)/", "", trial_name)) |>
  left_join(ed, by = "exp_id")
```


### Exclusions

Investigate distribution of RTs

```{r}
ggplot(exp_long, aes(x = rt)) +
  geom_histogram(bins = 100) +
  scale_x_log10(breaks = 10^(1:6),
                labels = c("10ms", "100ms", "1s", "10s", "100s", "1000s"))
```

And median RTs for peole who did at least 100 trials.

```{r}
med_rt_100 <- exp_long |>
  summarise(med_rt = median(rt), 
            n = n(),
            .by = session_id) |>
  filter(n > 100)

ggplot(med_rt_100, aes(x = med_rt)) +
  geom_histogram(bins = 100) +
  scale_x_continuous(breaks = seq(0, 10000, 500)) +
  labs(x = "Median Reaction Time (ms)")
```


Set RT median cutoff at 1% quantile?

```{r}
rt_cutoff <- quantile(med_rt_100$med_rt, probs = 0.01)
```


```{r}
# Calculate median RT and number of trials completed

ed <- select(exp_data, exp_id, trials)

rt <- exp_long |>
  summarise(start = min(dt),
            median_rt = median(rt), 
            mfvp = (table(dv) |> sort() |> tail(1) |> as.vector())/n(),
            n = n(),
            .by = c(session_id, exp_id)) |>
  left_join(ed, by = "exp_id")
```



```{r}
# Check the attention checks
checks <- exp_long |>
  select(session_id:dv) |>
  filter(grepl("check", trial_name)) |>
  mutate(check_type = sub("check_[a-z0-9-]+_", "", trial_name),
         check_type = ifelse(exp_id == 1400, substr(check_type, 4, 6), check_type)) |>
  summarise(checks_passed = mean(check_type == dv), 
            .by = c("session_id", "exp_id"))
```



```{r}
# Check try question
try <- quest_raw |>
  filter(q_name == "try") |>
  select(session_id, try = dv) |>
  unique()
```


```{r}
# combine to determine who gets excluded
start_date <- "2025-05-06"

to_exclude <- rt |>
  left_join(checks, by = c("session_id", "exp_id")) |>
  left_join(try, by = "session_id") |>
  filter(
    n != trials |
    interval(start, start_date) |> as.numeric("days") > 0 |
    is.na(checks_passed) |
    checks_passed < 5/7 |
    mfvp > 0.9 |
    median_rt < rt_cutoff |
    try != 2  
  ) 
```


```{r}
to_exclude |>
  mutate(reason = case_when(
    n < trials ~ "did not finish trials",
    n > trials ~ "too many trials",
    interval(start, start_date) |> as.numeric("days") > 0 ~ "date",
    checks_passed < 5/7 ~ "passed < 5/7 checks",
    mfvp > .9 ~ ">90% same response",
    median_rt < rt_cutoff ~ paste0("median RT < ", round(rt_cutoff)),
    try == 1 ~ "did not try",
    .default = "other")) |>
  count(reason, sort = TRUE)
  
```

Exclude `r nrow(to_exclude)` people and remove attention checks

```{r}
exp <- anti_join(exp_long, to_exclude, by = c("session_id", "exp_id")) |>
  filter(!grepl("check_", trial_name))
```

```{r}
endtimes <- quest_raw |>
  summarise(end = max(endtime), .by = c(session_id))

times <- exp |>
  summarise(start = min(dt), .by = c(session_id)) |>
  left_join(endtimes, by = "session_id") |>
  mutate(duration = interval(start, end) |> as.numeric("minutes"))
```

Number of remaining participants per study

There are `r unique(exp$session_id) |> length()` included participants.


```{r}
exp |>
  summarise(.by = c(exp, session_id)) |>
  count(exp)
```

## Demographics

```{r}
quest <- anti_join(quest_raw, to_exclude, by = c("session_id")) |>
  select(session_id, q_name, dv, endtime) |>
  unique() |>
  pivot_wider(names_from = q_name, values_from = dv) |>
  mutate(age = as.integer(age))
```

### Age and Gender

```{r}
ggplot(quest, aes(x = age, fill = gender)) +
  geom_histogram(binwidth = 1) +
  scale_fill_manual(values = c("hotpink", "lightblue", "orchid"))
```

### Residence

```{r}
count(quest, residence, sort = TRUE)
```

### Ethnicity

```{r}
quest |>
  mutate(ethnicity = tolower(ethnicity)) |>
  count(ethnicity, sort = TRUE)
```

### Devices

```{r}
count(quest, device, sort = TRUE)
```



## Plots

```{r}
rainbow <- c("firebrick", "darkorange", "goldenrod", "darkgreen", "dodgerblue3", "darkorchid4")
```

### Standardised Neutral Ratings

```{r}
exp_levels <- c("attractive", "trustworthy", "dominant", 
                "memorable", "gender-typical")

exp |>
  filter(exp_id %in% 1377:1382) |>
  mutate(dv = as.integer(dv),
         exp = factor(exp, exp_levels)) |>
  ggplot(aes(x = dv, fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 3, axes = "all_x", drop = FALSE) +
  labs(title = "Standardised Neutral Ratings",
       x = "") +
  scale_x_continuous(breaks = 1:7) +
  scale_fill_manual(values = rainbow, drop = FALSE)
```

```{r}
# function to create heatmap visualisations
heatmap <- function(id, label) {
  exp |>
    filter(exp %in% id) |>
    separate(trial_name, c("lab", "id"), extra = "drop") |>
    count(lab, id, dv) |>
    ggplot(aes(x = dv, y = id, fill = n)) +
    geom_tile() +
    facet_wrap(~lab) +
    scale_fill_viridis_c() +
    labs(x = label, y = NULL, 
         title = paste(label, "Ratings")) +
    theme(legend.position = "none", 
          axis.text.x = element_text(angle = 90))
}
```


#### Attractiveness

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("attractive", "Attractiveness")
```

#### Trustworthiness

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("trustworthy", "Trustworthiness")
```

#### Dominance

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("dominant", "Dominance")
```

#### Memorableness

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("memorable", "Memorableness")
 
```

#### Gender Typicality

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("gender-typical", "Gender Typicality")
```





### Unstandardised Neutral Ratings

```{r}
#| fig-height: 2.5
#| fig-width: 8
exp_labels <- c("attractive", "trustworthy", "dominant")
exp_levels <- paste(exp_labels, "(unstd)")

exp |>
  filter(exp_id %in% 1397:1399) |>
  mutate(dv = as.integer(dv), 
         exp = factor(exp, exp_levels, exp_labels)) |>
  ggplot(aes(x = dv, fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 5, drop = FALSE) +
  labs(title = "Unstandardised Neutral Ratings",
       x = "") +
  scale_x_continuous(breaks = 1:7) +
  scale_fill_manual(values = rainbow, drop = FALSE)
```

#### Attractiveness (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("attractive (unstd)", "Attractiveness (Unstandardised)")
```

#### Trustworthiness (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("trustworthy (unstd)", "Trustworthiness (Unstandardised)")
```

#### Dominance (Unstandardised)

```{r}
#| fig-height: 10
#| fig-width: 10

heatmap("dominant (unstd)", "Dominance (Unstandardised)")
```

### Emotion Ratings

```{r}
dv_levels <- c("anger", "disgust", "fear", 
                "happiness", "sadness", "surprise", "other")
emo_levels <- c("ang", "dis", "fea", "hap", "sad", "sur")
emo_labels <- paste(dv_levels[1:6], "faces")

exp |>
  filter(exp_id %in% c(1384, 1401:1405)) |>
  separate(trial_name, c("lab", "model", "type", "emo", "view")) |>
  mutate(dv = factor(dv, dv_levels),
         emo = factor(emo, emo_levels, emo_labels)) |>
  ggplot(aes(x = dv, fill = dv)) +
  geom_point(aes(x = x, colour = I(fill), fill = I(fill)), 
             data.frame(emo = factor(emo_levels, emo_levels, emo_labels),
                        x = 1:6,
                        fill = rainbow),
             size = 6.5, y = -60, shape = 18, show.legend = FALSE) +
  geom_bar(color = "transparent") +
  facet_wrap(~emo, axes = "all_x", drop = FALSE) +
  scale_x_discrete(labels = c("A", "D", "F", "H", "S", "U", "O")) +
  scale_fill_manual(values = c(rainbow, "grey"), drop = FALSE) +
  labs(title = "Emotion Ratings",
       x = "",
       fill = "Rated Emotion") +
  coord_cartesian(clip="off") +
  theme(axis.ticks.x = element_blank())
```

### Emotion Intensity Ratings

```{r}
exp_levels <- c("anger", "disgust", "fear", 
                "happiness", "sadness", "surprise")
               
exp |>
  filter(exp_id %in% 1385:1390) |>
  mutate(dv = as.integer(dv),
         exp = factor(exp, exp_levels)) |>
  ggplot(aes(x = dv, fill = exp)) +
  geom_histogram(binwidth = 1, show.legend = FALSE, color = "black") +
  facet_wrap(~exp, ncol = 3, axes = "all_x", drop = FALSE) +
  labs(title = "Emotion Intensity Ratings",
       x = "") +
  scale_fill_manual(values = rainbow, drop = FALSE) +
  scale_x_continuous(breaks = 1:7)
```

### Age Ratings

```{r}
dv_levels <- seq(20, 85, 5)
dv_labels <- paste(dv_levels-4, "-", dv_levels )
dv_labels[14] <- "81+"

exp |>
  filter(exp_id %in% 1400) |>
  mutate(dv = factor(dv, dv_levels, dv_labels)) |>
  ggplot(aes(x = dv)) +
  geom_bar(color = "black", fill = "white") +
  scale_x_discrete(drop = FALSE) +
  labs(title = "Age Ratings",
       x = "")
```

```{r}
# exp |>
#   filter(exp_id %in% 1400) |>
#   mutate(dv = as.numeric(dv) - 2.5) |>
#   summarise(age = mean(dv), age_sd = sd(dv), .by = trial_name)
```

```{r}
#| fig-height: 10
#| fig-width: 10

exp |>
  filter(exp_id %in% 1400) |>
  mutate(dv = factor(dv, dv_levels, dv_labels)) |>
  mutate(trial_name = gsub("_std_neu_0", "", trial_name)) |>
  separate(trial_name, c("lab", "id")) |>
  count(lab, id, dv) |>
  ggplot(aes(x = dv, y = id, fill = n)) +
  geom_tile() +
  facet_wrap(~lab) +
  scale_fill_viridis_c() +
  labs(x = "Age", y = NULL) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 90))
  
```



