---
title: "ManyFaces pilot ratings"
subtitle: "Data cleaning, exclusions, reliability analyses"
format: 
  html:
    code-fold: true
    df-print: paged
    embed-resources: true
    fig-height: 5
    fig-width: 8
    toc: true
execute: 
  error: true
---

```{r}
#| message: false
library(psych) # ICC, alpha, omega
library(furrr) # run code parallel
library(tidyverse)
```

```{r}
# Load custom functions
source("data/custom_functions.R")
```

## Project Structure

Read in the full project structure from the project file to map experiment names to numbers.

```{r}
proj <- jsonlite::read_json("data/project_1136_structure.json")

exp_data <- purrr::map_df(proj, \(comp) {
  if (comp$component_type == "exp") {
     data.frame(
       exp_id = comp$id,
       name = comp$name,
       res_name = comp$res_name,
       instructions = comp$instructions,
       question = comp$question,
       exptype = comp$exptype,
       trial_order = comp$trial_order,
       total_stim = comp$total_stim,
       random_stim = comp$random_stim,
       trials = length(comp$trial),
       stim = length(comp$stim)
     )
  } else {
    NULL
  }
}) |>
  mutate(exp = sub("ManyFaces? Pilot Ratings: ", "", res_name) |> trimws())


trial_data <- purrr::map_df(proj, \(comp) {
  if (comp$component_type == "exp") {
    purrr::map_df(comp$trial, \(trial) {
      data.frame(
        exp_id = comp$id,
        n = trial$trial_n,
        name = trial$name, 
        img_id = trial$center_img,
        img_path = comp$stimuli[[as.character(trial$center_img)]]
      )
    })
  } else {
    NULL
  }
}) |>
  mutate(name = sub("^(manyfaces|attention_checks)/", "", name))
```

```{r}
exp_data |>
  select(exp_id, exp, question, trials) |>
  arrange(exp)
```

## Data Prep

### Raw Data

This workflow requires the data-raw directory, which is not shared on github.

This is the SQL for downloading the data from Experimentum. We need to download in chunks of 50000 rows to avoid file download limits on the site (not needed if downloading directly from SQL).

``` sql
SELECT 
  session.id as session_id, project_id, exp.res_name as exp_name, exp_id, 
  session.user_id, user.sex as user_sex, user.status as user_status,
  ROUND(DATEDIFF(ed.dt, REPLACE(birthday, "-00","-01"))/365.25, 1) AS user_age,
  trial.name as trial_name,
  trial_n,
  `order`,
  dv,
  rt,
  ed.side,
  ed.dt
FROM session 
  LEFT JOIN user USING (user_id)
  LEFT JOIN exp_data AS ed ON ed.session_id = session.id
  LEFT JOIN exp ON exp.id = ed.exp_id
  LEFT JOIN trial USING (exp_id, trial_n)
WHERE session.project_id = 1136
  AND user.status IN ("guest", "registered")
  AND exp_id IN (1384, 1400, 1399, 1398, 1401, 1402, 1403, 
                 1404, 1405, 1397, 1390, 1389, 1388, 1387, 
                 1386, 1385, 1382, 1381, 1380, 1379, 1377)
LIMIT 50000
OFFSET 0
```

```{r}
#| eval: false

# combine multiple downloads into one file
exp_raw <- list.files("data-raw/exp", full.names = TRUE) |> 
  read_csv(show_col_types = FALSE) |>
  unique() |>
  filter(user_status %in% c("guest", "registered"))

write_csv(exp_raw, paste0("data-raw/ManyFaces-Pilot-Ratings-exps_", Sys.Date(), ".csv"))

# get most recent files
exp_file <- list.files("data-raw", "ManyFaces-Pilot-Ratings-exps",
                       full.names = TRUE) |>
  sort(decreasing = TRUE) |>
  pluck(1)

exp_raw <- read_csv(exp_file, show_col_types = FALSE) |>
  filter(user_status %in% c("guest", "registered")) |>
  unique()

# get most recent files
quest_file <- list.files("data-raw", "ManyFaces-Pilot-Ratings-quests", 
                       full.names = TRUE) |>
  sort(decreasing = TRUE) |>
  pluck(1)

quest_raw <- read_csv(quest_file, show_col_types = FALSE) |>
  filter(user_status %in% c("guest", "registered")) |>
  unique()

# write to the data directory
write_csv(exp_raw, "data/manyfaces-pilot-exp.csv")
write_csv(quest_raw, "data/manyfaces-pilot-quest.csv")
```

### Reshape Data

```{r}
exp_raw <- read_csv("data/manyfaces-pilot-exp.csv", show_col_types = FALSE)
quest_raw <- read_csv("data/manyfaces-pilot-quest.csv", show_col_types = FALSE)
```

```{r}
ed <- exp_data |>
  select(exp_id, exp, trials) |> 
  rename(trials_total = trials)

exp_long_raw <- exp_raw |>
  select(session_id, exp_id, trial_name, dv, rt, dt) |>
  unique() |>
  mutate(trial_name = sub("^(manyfaces|attention_checks)/", "", trial_name)) |>
  left_join(ed, by = "exp_id")
```

### Preliminary data cleaning

**Remove remove duplicate trials and incomplete data**

```{r}
#| echo: false
initial_participants <- exp_long_raw |> 
  count(session_id) |> 
  count() |> 
  pull(n)

initial_completions <- exp_long_raw |> 
  count(session_id, exp_id) |> 
  count() |> 
  pull(n)
```

We start with `r initial_participants` participants completing a total of `r initial_completions` experiments.

Some participants completed more than the maximum amount of trials in an experiment. Duplicate trials were removed by only retaining participants' initial ratings of a particular stimulus.

```{r}
dupl_trials_rm <- exp_long_raw |>
  arrange(session_id, exp_id, trial_name, trials_total, dt) |> 
  distinct(session_id, exp_id, trial_name, trials_total, .keep_all = TRUE)
```

```{r}
incomplete <- dupl_trials_rm |> 
  count(session_id, exp_id, trials_total) |> 
  filter(n < trials_total)

complete <- dupl_trials_rm |> 
  filter(!session_id %in% incomplete$session_id)
```

`r unique(incomplete$session_id) |> length()` participants did not complete all trials for a total of `r nrow(incomplete)` experiments.

As can be seen from discrepancy between number of participants and number of completed experiments, some participants completed more than one experiment. If participants completed more than one experiment, only data from the first was retained.

```{r}
dupl_exp_ids <- complete |> 
  # Mark participants that participated twice and establish which exp they did first
  group_by(session_id, exp_id) |> 
  summarise(first_dt = min(dt), .groups = "drop") |>
  group_by(session_id) |> 
  mutate(n_exp = n()) |>
  mutate(first_exp = exp_id[which.min(first_dt)]) |>
  ungroup() |> 
  filter(n_exp > 1 & exp_id != first_exp)
```

```{r}
exp_long <- complete |> 
   anti_join(dupl_exp_ids, by = c("session_id", "exp_id" = "first_exp"))
```

```{r}
#| echo: false
prelim_participants <- exp_long |> 
  count(session_id) |> 
  count() |> 
  pull(n)
```

Prior to pre-registered exclusions, preliminary sample thus consisted of `r prelim_participants` participants.

### Pre-registered exclusions

#### Overly consistent responses

Participants were excluded based on overly consistent responding, i.e. if they responded to at least 90% of trials identically.

```{r}
overly_consistent <- exp_long |> 
  summarise(
    same_pcnt = max(tabulate(match(dv, unique(dv)))) / n(),
    .by = c(session_id, exp_id)) |> 
  dplyr::select(session_id, exp_id, same_pcnt) |> 
  filter(same_pcnt >= 0.90)
```

#### Overly fast responses

Participants were excluded based on overly fast responding, i.e. if their median reaction time fell below the 1st percentile of the overall distribution of median reaction times.

```{r}
med_rt <- exp_long |>
  summarise(med_rt = median(rt),
            .by = c(session_id, exp_id))

overly_fast <- med_rt |> 
  filter(med_rt < quantile(med_rt, probs = 0.01))
```

```{r}
med_rt |> 
  ggplot(aes(x = med_rt)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = quantile(med_rt$med_rt, probs = 0.01), 
             color = "red", linetype = "dashed", linewidth = 1) +
  scale_x_continuous(breaks = seq(0, 10000, 500)) +
  labs(x = "Median Reaction Time (ms)") +
  theme_bw()
```

#### Self-reported honesty check

Participants were excluded based on self-reported honesty check, i.e. if they responded not taking the study seriously vs. taking it authentically

```{r}
honesty_check_failed <- quest_raw |>
  filter(q_name == "try") |>
  select(session_id, honesty_check = dv) |>
  unique() |> 
  filter(honesty_check != 2) |> 
  mutate(exp_id = NA) |>
  relocate(exp_id, .after = session_id)
```

#### Attention checks

Participants were excluded if they failed two or more attention checks

```{r}
attn_checks_failed <- exp_long |>
  select(session_id:dv) |>
  filter(grepl("check", trial_name)) |>
  mutate(check_type = sub("check_[a-z0-9-]+_", "", trial_name),
         check_type = ifelse(exp_id == 1400, substr(check_type, 4, 6), check_type)) |>
  summarise(attn_checks_passed = mean(check_type == dv), 
            .by = c("session_id", "exp_id")) |> 
  filter(attn_checks_passed < 5/7)
```

#### Overview of exclusions

```{r}
exclusions <- overly_consistent |> 
  full_join(overly_fast, by = c("session_id", "exp_id")) |> 
  full_join(honesty_check_failed, by = c("session_id", "exp_id")) |> 
  full_join(attn_checks_failed, by = c("session_id", "exp_id"))
```

```{r}
exclusions |> 
  summarise(
    `>90% same response` = sum(!is.na(same_pcnt)),
    `RT below cutoff` = sum(!is.na(med_rt)),
    `Failed honesty check` = sum(!is.na(honesty_check)),
    `Failed 2 or more of 7 attention checks` = sum(!is.na(attn_checks_passed)),
    "Total exclusions" = n()
  ) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Reason for exclusion",
    values_to = "Number of participants"
  )
```

Exclude `r nrow(exclusions)` participants and remove attention checks from data.

```{r}
data_exp <- exp_long |> 
  anti_join(exclusions, by = c("session_id", "exp_id")) |>
  filter(!grepl("check_", trial_name))
```

After exclusions, there were `r unique(data_exp$session_id) |> length()` participants.

Number of participants per study:

```{r}
data_exp |>
  summarise(.by = c(exp, session_id)) |>
  count(exp)
```

```{r}
endtimes <- quest_raw |>
  summarise(end = max(endtime), .by = c(session_id))

times <- data_exp |>
  summarise(start = min(dt), .by = c(session_id)) |>
  left_join(endtimes, by = "session_id") |>
  mutate(duration = interval(start, end) |> as.numeric("minutes"))
```

## Analyses

### Interrater agreement

```{r}
data_trait <- data_exp |>
  filter(exp == "attractive" | exp == "dominant"  | exp == "trustworthy"  | exp == "gender-typical"  | exp == "memorable") |>
  mutate(dv = as.numeric(dv))
  
data_trait_unstd <- data_exp |>
  filter(grepl("unstd", exp)) |>
  mutate(dv = as.numeric(dv))

data_emotion_int <- data_exp |>
  filter(exp == "anger" | exp == "happiness"  | exp == "disgust"  | exp == "surprise"  | exp == "sadness" | exp == "fear") |>
  mutate(dv = as.numeric(dv))

data_emotion_cat <- data_exp |> 
  # filter out emotion categorisation tasks
  filter(grepl("^em", exp))

```

#### ICC

##### Trait ratings

**Standardised**
```{r}
data_trait |> 
  group_by(exp) |> 
  group_map(calc_icc) |> 
  bind_rows()
```

**Unstandardised**
```{r}
data_trait_unstd |> 
  group_by(exp) |> 
  group_map(calc_icc) |> 
  bind_rows()
```

###### ICC "corridors of stability"

At what sample size (if at all) do ratings reach acceptable reliability of .75 to .90?

```{r}
#| warning: false
#| message: false
# Set seed for reproducibility
set.seed(123)

# Set a stable parallel plan
future::plan(multisession, workers = 3)

# Get unique experiments
experiments <- unique(data_trait$exp)

# Run the function in parallel across experiments
corridor_icc_list <- furrr::future_map(
  experiments,
  function(exp_name) {
    
    data_subset <- data_trait |>
      filter(exp == exp_name)
    
    calc_icc_cos(data = data_subset, 
           exp = exp_name, 
           n_raters_seq = seq(10, 100, 10), 
           n_iter = 50)
  },
  .options = furrr::furrr_options(seed = TRUE)
)

# Combine all results into a single tibble
corridor_icc_summary <- list_rbind(corridor_icc_list)

# Reset plan to sequential
future::plan(sequential)
```

```{r}
#| echo: false
icc2k_corridor_summary <- corridor_icc_summary |> 
  group_by(experiment, n_raters_sampled) |> 
  summarise(
    ICC2_k_median = median(`ICC(2,k)`, na.rm = TRUE),
    ICC2_k_low = quantile(`ICC(2,k)`, 0.025, na.rm = TRUE),
    ICC2_k_high = quantile(`ICC(2,k)`, 0.975, na.rm = TRUE),
    .groups = "drop"
  )

icc2k_corridor_summary |> 
  ggplot(aes(x = n_raters_sampled)) +
  geom_ribbon(aes(ymin = ICC2_k_low, ymax = ICC2_k_high), 
              alpha = 0.15, fill = "grey60") +
  # Median ICC(2,k)
  geom_line(aes(y = ICC2_k_median), linewidth = 0.8, colour = "black") +
  # Thresholds: 0.75 grey, 0.9 black
  geom_hline(yintercept = 0.75, linetype = "dashed", colour = "grey40") +
  geom_hline(yintercept = 0.9, linetype = "dashed", colour = "black") +
  facet_wrap(~ experiment) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    x = "Number of raters sampled", 
    y = "ICC(2,k)") +
  theme_minimal(base_size = 12) +
  theme(
    strip.text = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 12),
    plot.subtitle = element_text(size = 11)
  )
```

##### Emotion intensity

```{r}
data_emotion_int |> 
  group_by(exp) |> 
  group_map(calc_icc) |> 
  bind_rows()
```

#### Cronbach's alpha and McDonald's omega total

##### Trait ratings

**Standardised**
```{r}
#| warning: false
#| message: false
data_trait |> 
  group_by(exp) |> 
  group_map(calc_alpha_omega) |> 
  bind_rows()
```

Wee note re warnings (turned off in output):

-   "Number of categories should be increased...": occurs when some raters use only a subset of the response scale; that is not unexpected in subjective ratings (note that we did filter out raters with \<3 unique categories -- see also below)
-   "Some items were negatively correlated...": potentially reflects true inter-rater disagreement; reversing raters not appropriate in this context
-   "Omega_h... not meaningful": omega_h is undefined when computing a one-factor solution; omega_t remains valid and is the relevant statistic reported.

**Unstandardised**
```{r}
#| warning: false
#| message: false
data_trait_unstd |> 
  group_by(exp) |> 
  group_map(calc_alpha_omega) |> 
  bind_rows()
```

Four participants responded with only two values but were not caught by our "overly consistent filter". What do we want to do with them -- just drop them here, or exclude them more generally (as part of exclusions)?

```{r}
unique_values_check <- data_exp |> 
  # filter out emotion categorisation tasks
  filter(!grepl("^em", exp)) |>
  mutate(dv = as.numeric(dv))

unique_values_summary <- unique_values_check |> 
  group_by(exp) |> 
  group_map(~ {
    unique_counts <- check_unique_values(.x,
                                         col_map = list(p_id = "session_id",
                                                        trait = "exp",
                                                        stim_id = "trial_name",
                                                        rating = "dv"))
    tibble(
      exp = .y$exp,
      rater = names(unique_counts),
      n_unique = unique_counts
    )
  }) |> 
  bind_rows()

unique_values_summary |> filter(n_unique < 3)
```

##### Emotion intensity

```{r}
#| warning: false
#| message: false
data_emotion_int |> 
  group_by(exp) |> 
  group_map(calc_alpha_omega) |> 
  bind_rows()
```

### Point of stability (Hehman et al., 2018)

##### Trait ratings

**Standardised**
```{r}
set.seed(123)

stability_stats_traits <- calc_stability_stats(data = data_trait,
                                               N = 100,
                                               iterations = 300,
                                               ci_interval = 0.95, ci_method = "percentile",
                                               cos_threshold = 0.5,
                                               save_means = FALSE, # only save if actually needed (huge)
                                               col_map = list(trait = "exp",
                                                              stim_id = "trial_name",
                                                              rating = "dv")) 

# Summarise CI data: plotting resampled means is too much for my poor laptop
summary_ci_traits <- stability_stats_traits$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```

Plot
```{r}
# POS labels
stability_stats_traits$pos$label <- paste0("Point of stability\n(N = ", stability_stats_traits$pos$pos, ")")

summary_ci_traits |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_traits$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_traits$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```

Sanity check with Hehman et al. (20180) code

```{r}
source("data/resampling_hehmanetal.R")
```

```{r}
#| warning: false
#| message: false
# Trustworthy
trustworthy <- data_trait |> 
  filter(exp == "trustworthy")

trustworthy_rs <-RSSample(Rating = trustworthy$dv,
                          Stimulus = factor(trustworthy$trial_name))

# Our code N=38; Hehman et al. (2018) N=37
RSGetPointOfStability(trustworthy_rs$ci, 0.5, inarow = 1)

RSFigure(
  trustworthy_rs$data,
  trustworthy_rs$ci,
  threshold = .5,
  title = "",
  figPNGname = "",
  inarow = 1
)
```

```{r}
#| warning: false
#| message: false
# Memorable
memorable <- data_trait |> 
  filter(exp == "memorable")

memorable_rs <-RSSample(Rating = memorable$dv,
                        Stimulus = factor(memorable$trial_name))

# Our code N=45; Hehman et al. (2018) N=45
RSGetPointOfStability(memorable_rs$ci, 0.5, inarow = 1)

RSFigure(
  memorable_rs$data,
  memorable_rs$ci,
  threshold = .5,
  title = "",
  figPNGname = "",
  inarow = 1
)
```

**Unstandardised**
```{r}
set.seed(123)

stability_stats_traits_unstd <- calc_stability_stats(data = data_trait_unstd,
                                                     N = 100,
                                                     iterations = 300,
                                                     ci_interval = 0.95, ci_method = "percentile",
                                                     cos_threshold = 0.5,
                                                     save_means = FALSE,
                                                     col_map = list(trait = "exp",
                                                                    stim_id = "trial_name",
                                                                    rating = "dv")) 

summary_ci_traits_undstd <- stability_stats_traits_unstd$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```

Plot
```{r}
# POS labels
stability_stats_traits_unstd$pos$label <- paste0("Point of stability\n(N = ", stability_stats_traits_unstd$pos$pos, ")")

summary_ci_traits_undstd |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_traits_unstd$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_traits_unstd$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```

##### Emotion intensity
```{r}
set.seed(123)

stability_stats_emotion_int <- calc_stability_stats(data = data_emotion_int,
                                                    N = 100,
                                                    iterations = 300,
                                                    ci_interval = 0.95, ci_method = "percentile",
                                                    cos_threshold = 0.5,
                                                    save_means = FALSE,
                                                    col_map = list(trait = "exp",
                                                                   stim_id = "trial_name",
                                                                   rating = "dv")) 

summary_ci_emotion_int <- stability_stats_emotion_int$cis |> 
  group_by(exp, sample_size) |> 
  summarise(ul = mean(ul),
            ll = mean(ll),
            .groups = "drop")
```

Plot
```{r}
# POS labels
stability_stats_emotion_int$pos$label <- paste0("Point of stability\n(N = ", stability_stats_emotion_int$pos$pos, ")")

summary_ci_emotion_int |> 
  ggplot(aes(x = sample_size)) +
  geom_ribbon(aes(ymin = ll, ymax = ul, fill = exp),
              alpha = 0.5) +
  geom_vline(data = stability_stats_emotion_int$pos,
             aes(xintercept = pos),
             colour = "black",
             linetype = "solid",
             linewidth = .75) +
  geom_hline(yintercept = c(-0.5, 0.5),
             linetype = "dashed",
             colour = "grey30",
             linewidth = 0.5) +
  facet_wrap(~ exp) +
  xlab("Sample size (N)") +
  ylab("Centered mean ratings") +
  scale_fill_manual(values = paul_tol) +
  theme_minimal() +
  theme(panel.grid.major = element_line(colour = "#e8e8e8"),
        panel.grid.minor = element_line(colour = "#e8e8e8"),
        strip.text = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 12, face = "bold"),
        legend.position = "none") +
  geom_label(data = stability_stats_emotion_int$pos,
             aes(x = pos + 20, y = 1.6, label = label),
             fill = NA,   
             label.size = NA,
             colour = "black",    
             size = 3.5,
             inherit.aes = FALSE)
```
